---
title: "Vectors & Sequences"
author: "Hubert Baechli"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
  - \usepackage{xcolor}
  - \usepackage{titlesec}
  - \titleformat{\subsubsection}[block]{\normalfont\normalsize\bfseries\color{green}}{}{0em}{}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
GAP <- 1
```

# Programming exercise

Below are some exercises that should help you learn the most important basic built in functions and how to incorporate them into your own functions. This sheet focuses on vectors and how the easily generated.

## 1. by Sequences (easy)

The most challenging aspect of R is probably its flexibility. All these lines do exactly the same thing!

```{r}
vec_z <- 1:6
vec_z <- seq(1,6)
vec_z <- seq(1,6,1)
vec_z <- seq(from=1,to=6, by=1)
vec_z
```

***Try to generate this sequence***

```{r, echo = FALSE}
vec_z <- seq(from=1,to=6,by=2)
vec_z
```

## 2. by Repetitions (medium)

A second way to generate vectors is by using this:

```{r}
rep(x=5, times=4)
```

***Try to generate this sequence***

```{r, echo=FALSE}
vec_z2 <- rep(x=c(1, 3),times=4)
vec_z2
```

***And try now to generate this sequence with the same function***

Hint: you will need a other parameter then times (?rep)

```{r, echo=FALSE}
vec_z2 <- rep(x=c(1, 3),each=4)
vec_z2
```

## 3. Repetitions each(hard)

As you can see, the built-in functions are very flexible and can define many diferent vectors. I had to search longer to find a sequence that cannot be easily created. ;-)

***So create a function that generates this spezial sequence: (1-2-2-3-3-3-4 ...)***

*Hint: You will need a loop, and you have to concatenate the vectors together.*

```{r}
x <- c(2, 4) 
x <- c(x, c(3,4))
x
```

***Your result should do this at the end***

```{r, echo = FALSE}
rep_each <- function(x){
  vec <- c() 
  for (i in seq_along(x)) {
    vec <- c(vec, rep(x[i], x[i]))
  }
  return(vec)
}
```

```{r}
rep_each(x=c(1,3,2))
```
