---
title: "Kurves & Plots"
author: "Hubert Baechli"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
  - \usepackage{xcolor}
  - \usepackage{titlesec}
  - \titleformat{\subsubsection}[block]{\normalfont\normalsize\bfseries\color{green}}{}{0em}{}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
GAP <- 1
```

# Programming exercise

This sheet contains an exercise where you can apply and deepen the skills from E002 to E004

## 1. Calculating a Sine curve (E003)

To begin, we need a function that calculates the sine.

*Hint: implement at least this parameters, if you like you can also add offsets.*

```{r}
x <- pi
amp <- 2
freq <- 0.5

y <- amp * sin(freq * x)
y
```

***Your function now should do this***

```{r echo = FALSE}
calc_sin <- function(x, amp, freq, xoffset = 0, yoffset = 0){
  y <- amp * sin(freq * x + xoffset) + yoffset
  return(y)
}

```

```{r}
calc_sin(x = pi, amp = 2, freq = 0.5)
```

***and it should also work with a vector***

*Hint: sequences you learned in E004*

```{r}
x <- seq(from = 0, to = 4*pi, by = pi/4)
y <- calc_sin(x = x, amp = 2, freq = 0.5)
y
```

\newpage

## 2. Ploting a curve (combined)

Since the previous result is not easy to verify, let's create a plot of these results.

```{r}
plot(x, y, type = "l", col = "blue", lwd = 2, 
     xlab = "x", ylab = "y", main = "Sine curve")
```

You are already familiar with functions that return text or numbers. In R, however, it is also possible to write functions that generates directly plots.

```{r}
plot_sin <- function(x, GAP){
  y <- GAP # calc_sin
  title <- GAP
  return(plot(x, y, GAP))
}
```

\newpage

***You should got a function which does something like that***

*Hint: Use "E002"-skills for customizing your title*

```{r, echo=FALSE}
plot_sin <- function(x, 
                     amp, 
                     freq, 
                     xoffset = 0, 
                     yoffset = 0, 
                     color = "black"){
  y <- calc_sin(x = x, 
                amp = amp, 
                freq = freq, 
                xoffset = xoffset, 
                yoffset = yoffset)
  
  title <- paste("Sine curve (amplitude of", 
                 amp, 
                 "and frequency of", 
                 freq, 
                 ") in", 
                 color)
  
  return(plot(x, 
              y, 
              type = "l", 
              col = color, 
              lwd = 2, 
              xlab = "x", 
              ylab = "y", 
              main = title))
}
```

```{r}
x <- seq(from = 0, to = 4*pi, by = pi/25)
plot_sin(x = x, amp = 2, freq = 1, color = "green")
```

\newpage

## 3. Add a reverence curve (Outlook)

Once the basic building blocks are in place, it becomes easier to add further enhancements. Here, for example, is a reference curve added.

```{r, echo=FALSE}
plot_sin_rev <- function(x, 
                     amp, 
                     freq, 
                     xoffset = 0, 
                     yoffset = 0, 
                     color = "black"){
  
  y_rev <- calc_sin(x = x, 
                    amp = 1, 
                    freq = 1, 
                    xoffset = 0, 
                    yoffset = 0)
  
  plt <- plot_sin(x = x, 
                  amp = amp, 
                  freq = freq, 
                  xoffset = xoffset, 
                  yoffset = yoffset, 
                  color = color)
  
  plt <- plt + lines(x, y_rev, col = "black", lwd = 2, lty = 2)
}
```

```{r}
plot_sin_rev(x = x, amp = 2, freq = 2, color = "green")
```

\newpage

```{r}
plot_sin_rev(x = x, amp = 2, freq = 1, yoffset =1, color = "green")
```
