---
title: "E006: Optimizations"
author: "Hubert Baechli"
date: "2025-03-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bench)
GAP <- 1
```

::: callout-warning
This set of exercises is not fully ready and will be improved over the next few days. Do at your own risk ;)
:::

# Programming exercise

These exercises should help you write faster, more efficient code. So this exercises are inspired by both your LOGO code and mine (<https://hubiatgit.github.io/Starting_Github/LOGO/LOGO_Hubi.html>).

**Goal:** Improve the performance of the code below by creating more efficient functions.

::: callout-tip
For testing the speed you have to install the package "bench"

*see also <https://adv-r.hadley.nz/perf-measure.html>*
:::

**Data frame for testing the speed**

```{r}
hist <- data.frame(angle = seq(0, 360, 0.1))
hist$x <- 5 * cos(hist$angle / 180 * pi )
hist$y <- 5 * sin(hist$angle / 180 * pi )
hist$color <- sample(colors(), size = nrow(hist), replace = TRUE)
hist$color[sample(1:50, 1)] <- "red"

head(hist)
```

## 1. Constants

Doing things only ones if possible, means pre calculate often used numbers.

```{r}
new_pos <- function(x, y, hist) {
  new_x <- hist$x[[length(hist$x)]] + 2/3 * cos(hist$angle[nrow(hist)] * (pi / 180))
  new_y <- hist$y[[length(hist$y)]] + 2/3 * sin(hist$angle[nrow(hist)] * (pi / 180))
  return(c(new_x, new_y))
}

new_pos(10, 3, hist)
```

```{r, echo=FALSE}
new_pos_fast <- function(x, y, hist) {
  pos <- nrow(hist)
  rad <- hist$angle[pos] * (pi / 180)
  step <- 2/3
  new_x <- hist$x[pos] + step * cos(rad)
  new_y <- hist$y[pos] + step * sin(rad)
  return(c(new_x, new_y))
}
```

**Speed-test:**

```{r}
bench::mark(
  slow = new_pos(10, 3, hist),
  fast = new_pos_fast(10, 3, hist)
)
```

## 2. Early returns

Stop searching as soon as you've found what you're looking for. This idea can also be applied to a series of if-statements.

```{r}
find_first_red <- function(hist) {
 
  i <- 1
  res <- c() 
  
  while (i <= nrow(hist)) {
    if (hist$color[i] == "red") {
      res <- c(res, i)
    }
    i <- i + 1
  }
  return(max(min(res),0))
}

find_first_red(hist)
```

```{r, echo=FALSE}
find_first_red_fast <- function(hist) {
  
  i <- 1
  
  while (i <= nrow(hist)) {
    if (hist$color[i] == "red") {
      return(i)
    }
    i <- i + 1
  }
  return(0)
}

```

**Speed-test:**

```{r}
bench::mark(
  slow = find_first_red(hist),
  fast = find_first_red_fast(hist)
)
```

## 3. Omit Loops

Almost all functions in R can handle vector inputs. This means that loops aren’t truly avoided, but they are executed at the much faster underlying machine level.

```{r}
replot_hist <- function(hist) {

  plot(NULL,xlim = range(hist$x), ylim = range(hist$y), asp = 1)
  
  for (i in 2:nrow(hist)) {
    segments(x0 = hist$x[i - 1], 
             y0 = hist$y[i - 1],
             x1 = hist$x[i], 
             y1 = hist$y[i],
             col = hist$color[i])
  }
}
replot_hist(hist)
```

```{r, echo=FALSE}
replot_hist_fast <- function(hist) {
  pos <- nrow(hist)

  plot(NULL,xlim = range(hist$x), ylim = range(hist$y), asp = 1)
  
  segments(x0 = hist$x[-pos], 
           y0 = hist$y[-pos],
           x1 = hist$x[-1], 
           y1 = hist$y[-1],
           col = hist$color[-1])
}
```

**Speed-test:**

```{r}
bench::mark(
  slow = replot_hist(hist),
  fast = replot_hist_fast(hist)
)
```

## 4. Growing objects

Last but in my code possibly the worst: appending rows to an existing list or data frame using append(), c(old, new), or even rbind().

::: callout-tip
You will find better, faster ways to do this in <https://www.burns-stat.com/pages/Tutor/R_inferno.pdf>.
:::

```{r}
add_line <- function(x, y , hist) {

  last_angle <- hist$angle[length(hist$angle)]
  last_color <- hist$color[length(hist$color)]

  new_row <- data.frame(
    angle = last_angle,
    x = x,
    y = y,
    color = last_color
  )
  
  hist <- rbind(hist, new_row)
  hist
}

tail(add_line(3,10,hist))
```

```{r, echo=FALSE}
add_line_fast <- function(x, y , hist) {
  pos <- nrow(hist)
  
  hist[pos + 1, ] <- list(
    angle = hist$angle[pos],
    x = x,
    y = y,
    color = hist$color[pos]
  )
  hist
}
```

**Speed-test:**

```{r}
bench::mark(
  slow = add_line(3, 10, hist),
  fast = add_line_fast(3, 10, hist)
)
```

## 5. Logo

Now try to make your own Logo code faster. To begin, you might want to start by profiling your code, as described here: <https://adv-r.hadley.nz/perf-measure.html>

::: callout-tip
For this, you'll need a plain R script, which you can create like this:

```{r, eval = FALSE}
library(profvis)
library(bench)
library(knitr)

# convert your file to a R Script
purl("LOGO/LOGO_Hubi.qmd", output = "LOGO/LOGO_profiling.R")

# loading this file as a source
source("LOGO/LOGO_profiling.R")

# then run your logo inside profvis()
profvis(run_LOGO("CS SP [-pi/5] [-Sin(PI/2)/2] sPc [black]
                 REPEAT 20 [RepeaT 180 [FD 1 RT 2] RT 18]
                 SPEED 100
                 Home SpC [gREEN] StC [blue]
                 REPEaT 20.1 [rEPEAT 180 [fD 1 RT 2] RT 18]
                 UndoChunk
                 CS REPEAT 150 [ FD [n * 2]  RT 91 ]
                 SPEED 1 saVe [Logo04]"))
```
:::
